---
title: "FIGURES PhenoID: A framework to identify and explore cellular phenotypes in human organoids using a flow cytometry antibody panel"
output: pdf
---
Few libraries
```{r,include=FALSE}
require("flowCore") #Used for reading the data
require("flowTrans") #Used for transforming data
require("tidyverse")
require("stringr") #set of functions to manipulate strings type in order to have nice titles
require("rlist") #set of functions that allows to easely manipulate lists
require("reshape2") #visualization
require("scales") #scale colour intensity for visualization

# for clustering we use Rphenograph which needs to be installed directly from the github of JinmiaoChen Lab
if(!require(devtools)){
  install.packages("devtools") # If not already installed
}
devtools::install_github("JinmiaoChenLab/Rphenograph")
require("Rphenograph") #apply rphenograph clustering (Louvain method)
#Defines few functions to help
rename_markers<-function(flowset){#Defines a function to use marker names 
  copy_flowset=flowset[seq(along=flowset)]
  for (i in 1:length(copy_flowset)){
    marker.names=copy_flowset[[i]]@parameters@data$desc
    marker.names=lapply(marker.names,function(x){str_replace_all(x,"-","_")})
    colnames(copy_flowset[[i]]@exprs)<-unlist(lapply(marker.names, function(x){sapply(str_split(x,"_"),head,1)})) 
  }
  return(copy_flowset)
}
output_path="/home/alx/Documents/McGill_Neuro/Scripts/PhenoID/figures_paper/"
```

###FIGURE 1-b
Read in the data
```{r}
#Set the path of the file from the working directory
data_path="/home/alx/Documents/McGill_Neuro/Datasets/FACS_Paper/FIG2-A" 

#make a flowset object
exp=read.flowSet(path=data_path,transformation =FALSE ,emptyValue = FALSE,package="flowWorkspace")
# emptyValue = False doesn't consider empty rows

#select compensated channels, FCS and SSC that finishes by A (indicating the aea value) using regular expression
exp=fsApply(exp,function(x){x[ ,grepl("^[{'FJComp'}|{'FCS'}|{'SSC'}].*A$",colnames(exp))]})
#Still a flowset (fsApply -> apply a function to a flowset)

# if the data is located in a different space or you don't have the same type of fcs file exported (not complete) 
# don't run if the data is all the same folder and has all measures

data_path="/home/alx/Documents/McGill_Neuro/Datasets/FACS_Paper/FIG2"
exp1=read.flowSet(path=data_path,transformation =FALSE ,emptyValue = FALSE,package="flowWorkspace")
exp1=fsApply(exp1,function(x){x[ ,grepl("^[{'FJComp'}|{'FCS'}|{'SSC'}].*A$",colnames(exp1))]})#SELECT COMPENSED CHANNELS AND AEA VALUES
flowset=rbind2(exp,exp1) #FlowSet object
```

Facultative subsetting
```{r}
desired_size=1000
```
Set the number to the smallest fcs files (containing the smallest amount of cells)
```{r}
desired_size=min(fsApply(flowset,function(x){nrow(x@exprs)}))
```
Subsets the data
```{r}
set.seed(42) #Set a seed for reproducibility 
sf=sampleFilter(filterId = "SizeFilter", size =desired_size) #Creates a "filter" object to subset
flowset=fsApply(flowset,function(x){Subset(x,sf)}) #apply the filter on every flowframe
```

Check names
```{r}
sampleNames(flowset) #Prints the name of each flowframe inside the flowset. User can modify it in the following chunck
```

```{r}
sampleNames(flowset)<-c("3450c2","AIW2","AJG2","3450c1","3450c3","AIW1","AIW3","AJG1","AJG3") #directly modify the names inside the flowset
sampleNames(flowset)

#Transformation 
biexp  <- biexponentialTransform("biexp transform",a = 0.5, b = 1, c = 0.5, d = 1, f = 0, w = 0) #creates the transformation object
transformed_flowset <- transform(flowset, transformList(colnames(flowset), biexp)) #Apply the transformation
organo_transformed_flowset<-transformed_flowset#store for further usage
list_of_flowframes=fsApply(rename_markers(transformed_flowset),function(x){as.data.frame(x@exprs[,-c(1,2,9)])})#Makes a list of dataframes without FSC SSC and LiveDead markers
#hierarchical clustering for the heatmap
hclust=lapply(list_of_flowframes, function(x){hclust(dist(x,method="euclidean"),method="ward.D")$order})

#Reorder each list with the orders stored in hclust
for (i in names(list_of_flowframes)){
  list_of_flowframes[[i]]=list_of_flowframes[[i]][hclust[[i]],]
}

#Add order of cells for ploting the result
list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
for (index in seq_along(list_of_flowframes)){ #Iterates along the index
  list_of_flowframes[[index]]=list_of_flowframes[[index]] %>% #Using tidyverse for adding features
    mutate(Cells=as.factor(c(1:nrow(list_of_flowframes[[index]]))))
}
#Melt the result
mlf=melt(list_of_flowframes)
#Better names
colnames(mlf)<-c("Cells","marker","value","batch")
#order the melted dataframe in alphabetical order. We can impose a specific order by changing "levels"
mlf$marker=ordered(mlf$marker,levels=as.vector(unique(mlf$marker))[order(as.vector(unique(mlf$marker)))])

#plot
pdf(file =paste0(output_path,"mbos_heatmap.pdf"),11.7,8.3)#writes pdf (values are inches)
ggplot(mlf,aes(x=Cells,y=marker,fill=value))+geom_raster(size=0.5)+
  scale_fill_gradient2(low="cyan2", high ="darkorchid3")+
  scale_x_discrete(labels=mlf$batch,position = "top")+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(face="bold",size=12,angle=90),
        axis.text.y = element_text(face="bold",size=12),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank()
        )+
  facet_grid(~batch,scales="free",drop=FALSE)#heatmap
dev.off()

list_of_flowframes=fsApply(rename_markers(transformed_flowset),function(x){as.data.frame(x@exprs[,-9])})#Makes a list of dataframes
list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
for (index in seq_along(list_of_flowframes)){ #Iterates along the index
  list_of_flowframes[[index]]=list_of_flowframes[[index]] %>% #Using tidyverse for adding features
    mutate(Batch=list_names[index])
}
ds=list.rbind(list_of_flowframes)
ds$cell=as.factor(unlist(lapply(as.list(c(1:length(transformed_flowset))),function(x){c(1:nrow(transformed_flowset[[x]]@exprs))})))#add cell IDs 
write.csv(ds,file=paste0(output_path,"dataset_heatmap_mbos.csv"))

write.table(apply(as.data.frame(list.rbind(lapply(sampleNames(transformed_flowset),function(name){transformed_flowset[[name]]@description}))),2,as.character),file=paste0(output_path,"mbo_facs_description.csv"),col.names = TRUE)#Save descriptions of each flowframe in a csv file
```

###FIGURE 2 --- Same procedure
Reads in the data
```{r}
input_path="/home/alx/Documents/McGill_Neuro/Datasets/FACS_Paper/cell_line/"
flowset=read.flowSet(path=input_path,transformation = FALSE ,emptyValue = FALSE,package="flowCore") #Create a flowset object
flowset=fsApply(flowset,function(x){x[ ,grepl("^[{'FJComp'}|{'FCS'}|{'SSC'}].*A$",colnames(flowset))]})
```
Subsetting
```{r}
desired_size=5000
```
Set the number to the smallest fcs files (containing the smallest amount of cells)
```{r}
desired_size=min(fsApply(flowset,function(x){nrow(x@exprs)}))
```
Subsets the data
```{r}
set.seed(42) #Set a seed for reproducibility 
sf=sampleFilter(filterId = "SizeFilter", size =desired_size) #Creates a "filter" object to subset
flowset=fsApply(flowset,function(x){Subset(x,sf)}) #apply the filter on every flowframe
sampleNames(flowset)
```
#This chunck can be used for renaming the fcs files inside the flowset
```{r}
sampleNames(flowset)<-c("AstroSmall","neurons","NPC","oligo","astrocyte","IPSC","NPC2","oligo-2")
sampleNames(flowset)
```
###FIGURE 2-b
```{r}
#Transformation
biexp  <- biexponentialTransform("biexp transform",a = 0.5, b = 1, c = 0.5, d = 1, f = 0, w = 0) #creates the transformation object
transformed_flowset <- transform(flowset, transformList(colnames(flowset), biexp)) #Apply the transformation
list_of_flowframes=fsApply(rename_markers(transformed_flowset),function(x){as.data.frame(x@exprs[,-c(1,2,9)])})#Makes a list of dataframes

hclust=lapply(list_of_flowframes, function(x){hclust(dist(x,method="euclidean"),method="ward.D")$order})

#Reorder each list with the orders stored in hclust
for (i in names(list_of_flowframes)){
  list_of_flowframes[[i]]=list_of_flowframes[[i]][hclust[[i]],]
}

list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
for (index in seq_along(list_of_flowframes)){ #Iterates along the index
  list_of_flowframes[[index]]=list_of_flowframes[[index]] %>% #Using tidyverse for adding features
    mutate(
      Cells=as.factor(c(1:nrow(list_of_flowframes[[index]]))))
}

mlf=melt(list_of_flowframes)

colnames(mlf)<-c("Cells","marker","value","batch")
mlf=mlf[mlf$batch != "AstroSmall",]
mlf=mlf[mlf$batch != "NPC2",]

mlf$marker=ordered(mlf$marker,levels=as.vector(unique(mlf$marker))[order(as.vector(unique(mlf$marker)))])#order by marker

#plot
pdf(file =paste0(output_path,"cell_line_heatmap.pdf"),11.7,8.3)#writes pdf (values are inches)
ggplot(mlf,aes(x=Cells,y=marker,fill=value))+geom_raster(size=0.5)+
  scale_fill_gradient2(low="cyan2", high ="darkorchid3")+
  scale_x_discrete(labels=mlf$batch,position = "top")+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(face="bold",size=12,angle=90),
        axis.text.y = element_text(face="bold",size=12),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank()
        )+
  facet_grid(~batch,scales="free",drop=FALSE)#heatmap
dev.off()

list_of_flowframes=fsApply(rename_markers(transformed_flowset),function(x){as.data.frame(x@exprs[,-9])})#Makes a list of dataframes
list_names=names(list_of_flowframes)#extract names for not calling names() function at each loop
for (index in seq_along(list_of_flowframes)){ #Iterates along the index
  list_of_flowframes[[index]]=list_of_flowframes[[index]] %>% #Using tidyverse for adding features
    mutate(Batch=list_names[index])
}
ds=list.rbind(list_of_flowframes)
ds$cell=as.factor(unlist(lapply(as.list(c(1:length(transformed_flowset))),function(x){c(1:nrow(transformed_flowset[[x]]@exprs))})))#add cell IDs 
write.csv(ds,file=paste0(output_path,"dataset_heatmap_cell_lines.csv"))

write.table(apply(as.data.frame(list.rbind(lapply(sampleNames(transformed_flowset),function(name){transformed_flowset[[name]]@description}))),2,as.character),file=paste0(output_path,"cells_facs_description.csv"),col.names = TRUE)#Save descriptions of each flowframe in a csv file
```
Just to have some clustering results
```{r,echo=TRUE, results='hide'}
clustering=Rphenograph_flowset(organo_transformed_flowset,colstoignore = c(1,2,9),k=150, scale = FALSE)
```
###FIGURE 3-b
```{r}
binded=cbind(list.rbind(lapply(as.list(rename_markers(organo_transformed_flowset)@frames),function(x){as.data.frame(x@exprs[,-c(1,2,9)])})),clustering[[2]]$membership)
colnames(binded)[ncol(binded)]="cluster"
binded$cluster=as.factor(binded$cluster)

binded$Cells=as.factor(c(1:nrow(binded)))
mlb=melt(binded)
colnames(mlb)<-c("cluster","Cell","marker","value")

mlb$marker=ordered(mlb$marker,levels=as.vector(unique(mlb$marker))[order(as.vector(unique(mlb$marker)))])
pdf(file =paste0(output_path,"mbo_clustering_heatmap.pdf"),11.7,8.3)#writes pdf (values are inches)

ggplot(mlb,aes(x=Cell,y=marker,fill=value))+geom_raster(size=0.5)+
  scale_fill_gradient2(low="cyan2", high ="darkorchid3")+
  scale_x_discrete(labels=mlf$batch,position = "top")+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(face="bold",size=12,angle=90),
        axis.text.y = element_text(face="bold",size=12),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank()
        )+
  facet_grid(~cluster,scales="free",drop=FALSE)#heatmap
dev.off()

```

#Correlation
which file and which flowset
```{r}
expected_values=read.csv(file="/home/alx/Documents/McGill_Neuro/Datasets/FACS_Paper/ExpessionCellPhenotypeTable3.csv")
fs=organo_transformed_flowset
```

correlation
```{r}
#2 functions to help
minmax_normalize <- function(x){return((x- min(x)) /(max(x)-min(x)))}
compute_corr<-function(cell_profile,celltypes_profile){
  celltypes=celltypes_profile[,c(3:ncol(celltypes_profile))]#3 because the first two columns of the given file are expected to be "annotation" and "marker"
  res=as.vector(apply(celltypes,2,function(x){
    cor(cell_profile[which(!is.na(x))],x[which(!is.na(x))])}))#If a marker is "na" it won't be considered for computing the correlation coef
  names(res)<-colnames(celltypes)
  return(res)
}

fs=rename_markers(fs) #Nice marker names
facsdata=list.rbind(lapply(as.list(fs@frames),function(x){x=as.data.frame(x@exprs[,-9])})) #I had to remove the live-dead since it causes issues because it's not standardized
facsdata=as.data.frame(scale(as.matrix(facsdata),scale=TRUE,center=TRUE))#zscore by marker
facsdata=as.data.frame(apply(facsdata,2, minmax_normalize))#minmax normalize
rownames(expected_values)=toupper(expected_values$X)#Writes in upper case to make sure there is no mistake
colnames(facsdata)=toupper(colnames(facsdata))#same

intersect_markers=intersect(colnames(facsdata),rownames(expected_values))#Get common markers
intersect_markers=intersect_markers[order(intersect_markers)]#alphabetical order so there is no mistake
expected_values=expected_values[intersect_markers,]#ordering
facsdata=facsdata[,intersect_markers]#ordering
correlation_table=t(apply(facsdata,1,compute_corr,expected_values))#apply and store correlations by applying compute_corr function
assignments=t(apply(correlation_table,1,function(x){names(sort(x, decreasing=T))}))#get the list from the most resembling phenotype to the least resembling phenotype
print(head(correlation_table))
print(head(assignments))#Most resembling to less resembling 
```

